\section{Algorithm Analysis}

% \subsection{Proving Correctness of Algorithms}
\subsection{Stable Matching Problem}

\paragraph{Hospital's Stable Matching Problem}
Suppose there exists \(n\) doctors and hospital. Each hospital wants to
hire exactly one new doctor and the hospital ranks the priority of the
doctors they would prefer and each doctor also lists their preferences for
the hospitals they'd like to go to.

For a stable match, we would like to create an allocation where no group
would be happy to trade their allocations with each other.

\paragraph{Naive Solution}
The naive method computes all possible allocations and then picks one that
is stable.

This runs in \(O(n!)\) time which is not preferable.

\subparagraph{Fun Fact about Factorials - Sterling's Approximation}
\[n! \approx \left(\frac{n}{e}\right)^n\]

\paragraph{Gale - Shapely Algorithm - Assumptions}
\begin{itemize}
    \item Produces pairs in stages, with possible revisions.
    \item A hospital which has not been paired, will be called \textit{free}.
    \item Hospitals will offer jobs to doctors, who will decide whether to
    accept a job or not. Doctors may renege.
    \item All hospitals start off as free.
\end{itemize}

\paragraph{Gale - Shapely - Solution}

While there exists a free hospital which has not offered jobs to all doctors,
pick a free \(h\) and have it offer a job to the highest doctor \(d\) on its list to whom they have not offered a job yet.

If no one has offered \(d\) a job yet, they will always aspect the pair \(h, d\).

Otherwise, if they're already on the pair \(h', d\) then if \(h\) is a higher preference than \(h'\) the doctor will renege \(h'\) for \(h\) to form \((h, d)\). Otherwise, the hospital will seek another person to offer a job.

\paragraph{Proving Termination in \(n^2\)}
In all rounds of the Gale-Shapely algorithm, the hospital will offer a job to one doctor. They may offer a job to each doctor only once. Thus, the hospital may only make \(n\) offers. There are \(n\) hospitals each making \(n\) offers so, the algorithm must terminate in \(n^2\) rounds or less.

\paragraph{Proving Correctness of Gale-Shapely}

If the while loop has terminated with \(h\) still free, that implies that
\(h\) has offered a job to everyone. Since the only way to get declined is
if another, more appealing hospital has offered a job to other doctors.  

If there are no more doctors available, then \(n\) other hospitals must
have offered a job to \(n\) doctors. However this would imply there are
\(n + 1\) hospitals in total, which is a contradiction.

Hospitals will be happy as they start with their highest preference first and,
doctors will move up to accept their most preferred offer. Thus, there cannot exists a circumstance where both a hospital and doctor would like to swap.
